---
title: "Univariate Viz"
author: "Gabriel Donelly Higgins"
date: "April 12th, 2025"
---

## Research question:

What is the distribution of the length of titles for movies and tv shows on streaming platforms

## Step 1: Load necessary packages

```{r}
library(tidyverse)
library(ggthemes)
```


## Step 2: Import and clean / wrangle data

```{r}
messy_movies <- read_csv("../data/MoviesOnStreamingPlatforms.csv")

movies_len <- messy_movies |>
  mutate(title_len = str_length(Title)) |> # I am mutating a variable that counts the length of the Title variable
  select(title_len) # since the title_len variable is the only thing I need, I am only going to select that variable in this dataset I am creating
```


## Step 3: Put data into a plot

```{r}
#| fig-cap: "Histogram of the distribution of movies and TV shows by the total number of characters in their title"
#| fig-alt: "A histogram with an x-axis titled 'Total Characters in Title' and a y-axis titled 'Movies and TV Shows'. The bars on the histogram have a grey color. The x-axis has a range of 120, and the y-axis has a range of 2500. The bins are have a width of 5, and the bin with the most observations is the 11-15 bin, closely followed by the 6-10 bin. They each have slightly more than 2500 observations, while the only other bin that comes close is the 16-20 bin, with less than 1500 observations. The distribution of movies and TV shows in the bins is initially low, at slightly more than 1000 observations for the 1-5 bin, followed by the peaks of the 6-10 and 11-15 bins, and then decreasing exponentially across each bin. There is no movie or TV show with more than 60 characters in its title"

ggplot(movies_len, aes(x = title_len)) +
  geom_histogram(color = "cornsilk", # this helps to make the bars contrast against each other
                 binwidth = 5) + # this allows the data to be broken into classes, making it easier to explain and understand
  theme_solarized() +
  labs(x = "Total Characters in Title",
       y = "Movies and TV Shows",
       title = "Movies and TV Shows on Streaming Services Versus the Total Characters in Their Title")
```


## Conclusion:

Movies and TV shows on streaming services tends to offer less entertainment as the title of the entertainment more characters than 15. A plurality of the entertainment has 11-15 characters, sharply followed by 6-10. As one looks for movies and tv shows with titles longer than 15 characetrs, the number of observations decreases exponentially. There could be many reasons for this such as audiences' attention spans preferring shorter titles or writers simply coming up with shorter titles more often than very long ones. 


## References:
https://www.kaggle.com/datasets/ruchi798/movies-on-netflix-prime-video-hulu-and-disney (accessed: April 11th, 2025)
