---
title: "exam_review"
format: html
---
```{r}
library(tidyverse)
```
## The Max Number of Variables in a Viz is 7:

 - X-axis
 - Y-axis
 - Opacity
 - Color
 - Size
 - Shape
 - Facets
 - 3 dimensions

## Effective Viz Checklist:

 - Is graph effective?
  - right for data
  - appropriate for story
  - clear purpose
  - appropriate audience
  - unbiased / inclusive
  - not misleading

 - Is graph complete?
  - title
  - subtitle
  - X-axis label
  - X-axis marks
  - Y-axis label
  - Y-axis marks
  - legend title
  - legend labels
  - capition (author, date, source)
  - data labels
  - date annotation

 - Is graph accessible
  - color (fill, border, background, text)
  - alt-text
  - text size

## Steps to an Effective Viz:

 - Identify simple research questions
  - what do you want to understand about the variables or the relationships among them?
 - Start with the basic and work incrementally
  - Identify what variables you want to include in your plot and what structure these have (e.g: categorical,      quantitative, dates)
  - start simply, build a plot of just one of these variables, or the relationship between two of the variables
  - set up a plotting frame and add just one geometric layer at a time
  - start tweaking: add whatever new variables you want to examine
 - Ask your plot questions
  - what questions does your plot answer? What questions are left unanswered by your plot?
  - what new questions does your plot spark/inspire?
  - do you have the visualization tools to answer these questions, or might you learn more?
 - Focus
  - reporting a large number of visualizations can overwhelm the audience and obscure your conclusions.            Instead, pick out a focused yet comprehensive set of visualizations.

## Data in Spreadsheets is Called Tidy When:

 - Each row = a unit of observation
 - Each column = a measure of some variable of interest:
  - quantitative = numbers with units
  - categorical = discrete possibilities or categories
 - Each entry contains a single data value, ie, no analysis, summaries, footnotes, comments, etc. Only one        value per cell

## Bivariate / Multivariate Viz:

In bivariate viz there are two variables. One is an independent variable and one is a dependent variable. You are trying to explain the variability of one variable relative to the other. 

```{r}
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

ggplot(elections, aes(x = repub_pct_16, y = repub_pct_20)) +
  geom_point() +
  geom_smooth(method = "lm")
  labs(x = "Percent of Republican Vote in 2016", title = "Republican Vote by County in 2016 vs 2020", y = "Percent of Republican Support in 2020")
```

In multivariate viz, you just use more variables to show the relationship between them.

```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "Temperature at 9 am", title = "Weather observation in Australia",y = "Temperature at 3 pm", color = "Location")
```

