---
title: "Exam 2"
format: html
---

* this is a continuation of exam one

## Research question:

**What does the consumption of each food category in each country look like?**


## Import Necessary Data and Tools:

**Load in packages:**
```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
library(ggthemes)
```


**Load data:**
```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```


**Inspect Data:**
```{r}
str(fc)
```

```{r}
head(fc, 22)
```


**Inspect Country Variable:**
```{r}
fc |>
  group_by(country) |>
  summarize(sum(consumption)) |>
  str()
```

There are 130 observations, when you set the unit of observation to countries, so there are 130 countries in this dataset


**Inspect food_category variable:**
```{r}
fc |>
  distinct(food_category)
```

Here we have 11 distinct food categories, but it is better to rename them for the sake of making them easier to read in a viz

**Clean data:**
```{r}
fc_clean <- fc |>
  mutate(food_category = as.factor(food_category),
         food_category = fct_recode(food_category,
                                    "Lamb" = "Lamb & Goat",
                                    "Dairy" = "Milk - inc. cheese",
                                    "Wheat" = "Wheat and Wheat Products",
                                    "Nuts" = "Nuts inc. Peanut Butter"))
```

Double checking to make sure all variables are properly renamed
```{r}
fc_clean |>
  distinct(food_category)
```

## Sub-research Question 1: Which 5 countries consume the most food?

**Expected Dataframe:**
We want a dataframe in which countries are listed, followed by their total consumption


**Reshaping / Wrangling data:**
```{r}
country_consumption <- fc_clean |>
  group_by(country) |>
  summarize(total_consumption = sum(consumption))
  # pivot_wider(names_from = food_category, values_from = consumption) |>
  # mutate(total_consumption = Pork + Poultry + Beef + Lamb + Fish + Eggs + Dairy + Wheat + Rice + Soybeans + Nuts) |>
  # select(country, total_consumption)
```


**Visualizing Data:**
```{r}
country_consumption |>
  arrange(desc(total_consumption)) |>
  head(5) |>
  ggplot(aes(x = country, y = total_consumption)) +
    geom_col() +
    theme_solarized()
```


**Observations:**
Seemingly, the top consumers of food are exclusively European countries. Their geographic location varies from Eastern to Western, Baltics and Balkans and Scandinavian countries. They all consume roughly similar amounts of food, with Findland consuming the most.


## Sub-research Question 2: Which 5 countries consume the most of each food?

**Expected Dataframe:**
We want a dataframe where there is a variable for countries, a variable for the food type, and a variable for the consumption value. The food_category should correspond to the top 5 highest values in the consumption variable for its category. And the countries in the country variable will be the top 5 consumers of whatever food category is in the food variable.


**Reshaping / Wrangling Data:**
```{r}
top_food_consumers <- fc_clean |>
  select(-co2_emmission) |>
  group_by(food_category) |>
  slice_max(order_by = consumption, n = 5)
```


**Vizualizing Data:**
```{r}
ggplot(top_food_consumers, aes(x = food_category, y = consumption, fill = country)) +
  geom_col() +
  theme_solarized() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 5),
        axis.text.y = element_text(size = 5))
```


**Observations:**
Of all categories, dairy, wheat, and rice are the most consumed. The consumption of dairy, wheat, or rice is enough to outmatch the entire consumption of the top 5 countries combined in some categories like nuts or soybeans. What this means is that some of these categories are relatively easy to be a top 5 country in, because their consumption is very low anyway. 


## Sub-research Question 1: What does the consumption of each food look like?

**Expected Dataframe:**
We want a dataframe in which there is a country variable, a variable for food type, a corresponding variable for consumption, and a variable for the geometry of each country.


**Reshaping, Wrangling, and importing data:**

Find out which countries don't have a match in the geometry dataset
```{r}
fc_clean |>
  anti_join(ne_countries(), join_by(country == admin)) |>
  distinct(country)

st_set_geometry(ne_countries(), NULL) |>
  anti_join(fc_clean, join_by(admin == country)) |>
  distinct(admin)
```


Rename the countries so that they have a match and join the shapefile
```{r}
temp <- fc_clean |>
  select(-co2_emmission) |>
  group_by(food_category) |>
  mutate(consumption = (consumption - mean(consumption)) / sd(consumption)) |>
  mutate(country = fct_recode(country,
                              "United States of America" = "USA",
                              "Czechia" = "Czech Republic",
                              "Republic of Serbia" = "Serbia",
                              "eSwatini" = "Swaziland",
                              "North Macedonia" = "Macedonia",
                              "Mauritania" = "Mauritius",
                              "United Republic of Tanzania" = "Tanzania",
                              "Democratic Republic of the Congo" = "Congo",
                              "United Republic of Tanzania" = "Tanzania",
                              "The Bahamas" = "Bahamas",
                              "Taiwan" = "Taiwan. ROC"))

temp

global_consumption <- ne_countries() |>
  select(admin, geometry) |>
  left_join(temp, join_by(admin == country)) |>
  mutate(consumption = replace_na(consumption, 0)) |>
  pivot_wider(names_from = food_category, values_from = consumption) |>
  select(-`NA`) |>
  pivot_longer(cols = c(-geometry, -admin), names_to = "food_category", values_to = "consumption")

head(global_consumption)
```


**Vizualizing Data:**
```{r}
#| fig-height: 8
#| fig-width: 15
ggplot() +
  geom_sf(data = global_consumption,
          aes(fill = consumption)) +
  facet_wrap(~food_category) +
  coord_sf(crs = "+proj=robin") +
  theme_map() +
  theme(legend.position = c(0.8, 0.05))
```

**Observations:**
Beef is heavily consumed in many highly countries like the United States, Russia, and Brazil, but also less consumed around most of the world and in other large countries like China, India, and Mexico. Dairy is heavily consumed in smaller countries like Finland and Kazakhstan, with moderate consumption around Europe and North America, and generally less consumption worldwide. Eggs are heavily consumed in China, with moderate consumption across Europe and the Americas. Fish has low consumption worldwide, but Burma stands out as a high consumer. Lamb is heavily consumed in Kazakhstan, Algeria, and Australia, while the rest of the world has considerably less consumption. The largest consumers of nuts are Persia, Canada and North America, and Bolivia. Additionally, countries with more moderate consumption include China, Australia, and Turkey. Pork is highly consumed in Russia, Europe and North America compared to the rest of the world. Poultry products tend to be more consumed in the Americas, Australia, and Saudi Arabia, with less worldwide consumption elsewhere. Rice is highly consumed in East, Southeast, and South Asia. Soybeans are highly consumed in China, Brazil, Botswana, and Zimbabwe. And lastly, Wheat is highly consumed in Algeria, Russia, Egypt, Turkey, and Iran. It is otherwise less consumed elsewhere.


## Conclusion:

Most countries around the world have unique eating habits, but regional patterns to appear. For example, rice is heavily consumed in South, Southeast, and East Asia. The entirety of the Americas consumes very little lamb. Yet Kazakhstan, Australia, and Algeria (countries with large portions of lowly populated land) consume a lot of lamb. Similarly, beef is highly consumed in Brazil, the United States, and Australia. These countries all have lots of suitable land for cattle to graze. Dairy also seems to follow lactose tolerance, to an extent, with less consumption in much of Asia and Africa. However, throughout the Americas, where there are mixed and indigenous populations with higher rates of lactose intolerance, dairy still has a presence. Certain food categories tended to have a few outliers, like soybeans, may have had higher consumption in certain countries (like Botswana and Zimbabwe), but their worldwide consumption was so small that just the consumption of wheat among one of the top 5 wheat consumers was more than quadruple the consumption of all the top 5 soybean consumers combined. Surprisingly, the countries that had the highest total consumption were Albania, Finland, Lithuania, the Netherlands, and Sweden. These were all coastal European countries with less than one 100th of India's population. I expected to see larger countries, like Brazil, Japan, India, China, Russia, or the United States to be top consumers. What this may indicate is that consumption is not calculated as the absolute consumption of food in these countries, but perhaps it is done in some other way. Considering that the top consumers were all coastal, perhaps calculations were based on food moving through the country, and these European countries had ports with high traffic. Or perhaps data wasn't sufficiently collected in larger countries like India, China, or the United States. 












