---
title: "exam_review"
format: html
---
```{r}
install.packages(c("sf", "maps", "RColorBrewer", "gplots", "socviz", "leaflet", "devtools", "ggthemes"))
library(tidyverse)
library(sf)
library(maps)
library(mosaic)
library(gplots)
library(leaflet)
```
## The Max Number of Variables in a Viz is 7:

 - X-axis
 - Y-axis
 - Opacity
 - Color
 - Size
 - Shape
 - Facets
 - 3 dimensions

## Effective Viz Checklist:

 - Is graph effective?
  - right for data
  - appropriate for story
  - clear purpose
  - appropriate audience
  - unbiased / inclusive
  - not misleading

 - Is graph complete?
  - title
  - subtitle
  - X-axis label
  - X-axis marks
  - Y-axis label
  - Y-axis marks
  - legend title
  - legend labels
  - capition (author, date, source)
  - data labels
  - date annotation

 - Is graph accessible
  - color (fill, border, background, text)
  - alt-text
  - text size

## Steps to an Effective Viz:

 - Identify simple research questions
  - what do you want to understand about the variables or the relationships among them?
 - Start with the basic and work incrementally
  - Identify what variables you want to include in your plot and what structure these have (e.g: categorical,      quantitative, dates)
  - start simply, build a plot of just one of these variables, or the relationship between two of the variables
  - set up a plotting frame and add just one geometric layer at a time
  - start tweaking: add whatever new variables you want to examine
 - Ask your plot questions
  - what questions does your plot answer? What questions are left unanswered by your plot?
  - what new questions does your plot spark/inspire?
  - do you have the visualization tools to answer these questions, or might you learn more?
 - Focus
  - reporting a large number of visualizations can overwhelm the audience and obscure your conclusions.            Instead, pick out a focused yet comprehensive set of visualizations.

## Data in Spreadsheets is Called Tidy When:

 - Each row = a unit of observation
 - Each column = a measure of some variable of interest:
  - quantitative = numbers with units
  - categorical = discrete possibilities or categories
 - Each entry contains a single data value, ie, no analysis, summaries, footnotes, comments, etc. Only one        value per cell

## Bivariate / Multivariate Viz:

In bivariate viz there are two variables. One is an independent variable and one is a dependent variable. You are trying to explain the variability of one variable relative to the other. 

```{r}
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

ggplot(elections, aes(x = repub_pct_16, y = repub_pct_20)) +
  geom_point() +
  geom_smooth(method = "lm")
  labs(x = "Percent of Republican Vote in 2016", title = "Republican Vote by County in 2016 vs 2020", y = "Percent of Republican Support in 2020")
```

In multivariate viz, you just use more variables to show the relationship between them.

```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "Temperature at 9 am", title = "Weather observation in Australia",y = "Temperature at 3 pm", color = "Location")
```

## Spatial Viz:

Create a map widget by calling leaflet( ) and telling it the data to use
```{r}
fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")

leaflet(data = fave_places)
```

Add a base map using addTiles( ) or addProviderTiles( )
```{r}
leaflet(data = fave_places) |> 
  addTiles()
```

Add layers to the map using layer functions(addMarkers( ) or addPolygons( ))
```{r}
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers(lng = ~longitude, lat = ~latitude)
```
Print the map widget to display it


## The 3 Types of Spatial Viz:

Point map: Maps plotting individual observations
```{r}
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")

ggplot(starbucks, aes(y = Latitude, x = Longitude)) + 
  geom_point(size = 0.5)
```


Contour maps: Maps plotting the density of the distribution of observations
```{r}
starbucks_cma <- starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))

cma_boundaries <- ne_states(
  country = c("canada", "mexico", "united states of america"),
  returnclass = "sf")

ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_density_2d(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```


Choropleth maps: Maps plotting outcomes in different regions
```{r}
elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))

states_map <- map_data("state")

ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() +
  labs(fill = "% of Republican Support")
```


## Alt text and titles

Alt text is a written description of the viz that can be read by a screen reader
```{r}
#| fig-alt: "This is a density plot containing a numerical and a categorical variable. The numerical variable in the temperature at 3 pm in Celcius. The categorical variable is the location where the temperature was measured. The numerical variable, the temperature, is on the x-axis. The y-axis represents the density of observations. The density plot has created 3 different lumps, each color coded for the legend which matches them to their location. The Hobart lump rises from 0 degrees celcius to 12.5, where it plateus before falling again at 17.5 and becoming exponentially lower on the y-axis as it approaches 40 degrees celcius on the x-axis. The Uluru lump rises from 10 degrees celcius, and has a series of plataeus followed by elevations until it reaches 37.5 degrees where it begins its decline till it reaches 40 degrees celcius on the x-axis. The Wollongong lump rises rapidly at 12.5 degrees celcius, peaks at 20 degrees celcius, and then rapidly falls till 27.5 degrees celcius where it then has a slightly negative plateuatill 40 degrees celcius."

ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)")  
```

The title is a 1-sentence description of a plot
```{r}
#| fig-cap: "Density plots of 3pm temperatures in 3 Australian locations."

ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)")  
```

## Categorical plots

**1 variable:**

Bar chart (just counts observations)
```{r}
hikes <- read.csv("https://mac-stat.github.io/data/high_peaks.csv")

ggplot(hikes, aes(x = rating)) +
  geom_bar()
```

**2 variables:**

Bar charts:
  - stacked (heights vary, but each color (stripe) of the bar follows the Y-axis)
```{r}
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```
  - dodge (side by side)
```{r}
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "dodge")
```
  - proportional (all same height)
    - this loses information about the number of observations
```{r}
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill")
```

**3 variables:**

Facted wrapped bar charts:
  - facet wrapped stacked
```{r}
ggplot(weather, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar() +
  facet_wrap(~location)
```
  - facet wrapped dodge
```{r}
ggplot(weather, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "dodge") +
  facet_wrap(~location)
```
  - facet wrapped proportional
```{r}
ggplot(weather, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill") +
  facet_wrap(~location)
```

## Numerical Plots:

**1 variable:**

Histogram
```{r}
ggplot(weather, aes(x = maxtemp)) +
  geom_histogram(color = "white")
```

Box plot
```{r}
ggplot(weather, aes(y = maxtemp)) +
  geom_boxplot()
```

Density plot
```{r}
ggplot(weather, aes(x = maxtemp)) +
  geom_density()
```

**2 variables:**

Scatterplot
```{r}
ggplot(weather, aes(x = maxtemp, y = mintemp)) +
  geom_point()
```

Line plot
```{r}
ggplot(weather, aes(x = maxtemp, y = mintemp)) +
  geom_line()
```

**3 variables:**

Scatterplot
  - color
```{r}
ggplot(weather, aes(x = maxtemp, y = mintemp, color = windspeed9am)) +
  geom_point(alpha = 1)
```
  - proportional symbols
```{r}
ggplot(weather, aes(x = maxtemp, y = mintemp, size = rainfall)) +
  geom_point(alpha = 0.5) +
  scale_size_continuous(name = "rainfall", range = c(1, 5))
```
  - opacity
```{r}
ggplot(weather, aes(x = maxtemp, y = mintemp, alpha = windgustspeed)) +
  geom_point()
```

## Mixed Variable Plots:

**1 numerical + 1 categorical:**

Density plots:
  - faceted
```{r}
ggplot(weather, aes(x = windgustspeed)) +
  geom_density() +
  facet_wrap(~location)
```
  - color
```{r}
ggplot(weather, aes(x = windgustspeed, fill = location)) +
  geom_density(alpha = 0.5)
```

Box plot (faceted)
```{r}
ggplot(weather, aes(y = windgustspeed)) +
  geom_boxplot() +
  facet_wrap(~location)
```

Violin
```{r}
ggplot(weather, aes(y = windgustspeed, x = location)) +
  geom_violin()
```
Histograms:
  - faceted
```{r}
ggplot(weather, aes(x = windgustspeed)) +
  geom_histogram(color = "white") +
  facet_wrap(~location)
```
  - fill / color
```{r}
ggplot(weather, aes(x = windgustspeed, fill = location)) +
  geom_histogram(color = "white")
```

**1 numerical + 2 categorical:**

Heat map
```{r}
#| fig-height: 15
#| fig-width: 8
education <- read.csv("https://mac-stat.github.io/data/sat.csv")

plot_data <- education |> 
  column_to_rownames("State") |> 
  data.matrix() |> 
  scale()

heatmap.2(plot_data,
  dendrogram = "none",
  Rowv = NA, 
  scale = "column",
  keysize = 0.7, 
  density.info = "none",
  col = hcl.colors(256), 
  margins = c(10, 20),
  colsep = c(1:7), rowsep = (1:50), sepwidth = c(0.05, 0.05),
  sepcolor = "white", trace = "none"
)
```


**2 numerical + 1 categorical:**

Scatterplots:
  - faceted
```{r}
ggplot(weather, aes(x = mintemp, y = maxtemp)) +
  geom_point() +
  facet_wrap(~location)
```
  - fill / coor
```{r}
ggplot(weather, aes(x = mintemp, y = maxtemp, color = location)) +
  geom_point()
```
  - shape
```{r}
ggplot(weather, aes(x = mintemp, y = maxtemp, shape = location)) +
  geom_point(size = 1, alpha = 0.5)
```
  - labeled points
```{r}
ggplot(weather, aes(x = mintemp, y = maxtemp)) +
  geom_text(aes(label = windgustdir), alpha = 1, size = 2)
```

## Data manipulation

**Piping:**
Use the |> pipe operator. It makes the object on the left side pass to the function on the right to the get a new object.

**Types of reshaped data:**
- Aggregated data
  - Data that has had functions like group_by() or summarize() done to it. It gains aggregate information         about the observations but loses data on individual observations
- Raw data, reshaped
  - Data that is reshaped to maintain individual observations, but in a way that it can be properly               manipulated

**Unit of observation:**
The level at which data is collected and analyzed, by individual cases or groups.

For example, in a dataset of students, each student is the unit of observation. In a dataset of cities, each city is the unit of observation. In a datset of daily weather data, each day is the unit of observation

## Command List:

| Verb             | Outcome                                                                 |
|------------------|-------------------------------------------------------------------------|
| `arrange`        | Arrange the rows according to a variable                                |
| `filter`         | Filter out or obtain a subset of the rows                               |
| `select`         | Select a subset of the columns                                           |
| `mutate`         | Mutate or create a column                                               |
| `summarize`      | Calculate a numerical summary of a column                               |
| `group_by`       | Group the rows by a specified column                                    |
| `ungroup`        | Ungroups the previous `group_by`                                        |
| `count()`        | Counts the number of observations in a group or just the whole column   |
| `n()`            | Also counts the number of observations in a group, but must be used inside of `summarize` or `mutate` |
| `na.rm`          | Makes it so that the expressions you use in your command do not have the `NA` value |
| `as.factor`      | Use it in the `mutate` command to make a variable categorical           |
| `as.numeric`     | Use it in the `mutate` command to make a variable numeric               |
| `head()`         | Shows first rows of dataset                                             |
| `library(package)` | Loads package into program                                            |


**arrange:**

We have dataset weather
```{r}
head(weather)
```
And we use arrange(maxtemp) on it to order observations from lowest to highest value in maxtemp
```{r}
weather |>
  arrange(maxtemp) |>
  head()
```
Or arrange(desc(maxtemp)) to order observations from highest to lowest value in maxtemp
```{r}
weather |>
  arrange(desc(maxtemp)) |>
  head()
```

**filter:**

We can use filter(maxtemp < 45) to get observations with a maxtemp below 45
```{r}
weather |>
  filter(maxtemp < 45) |>
  head()
```

**select:**

We can use select(maxtemp, mintemp, location) to only select those columns
```{r}
weather |>
  select(maxtemp, mintemp, location) |>
  head()
```
Or we can use (-maxtemp) to select all columns except maxtemp
```{r}
weather |>
  select(-maxtemp) |>
  head()
```

**mutate:**

We can use mutate(double_maxtemp = maxtemp * 2) to create a new column
```{r}
weather |>
  mutate(double_maxtemp = maxtemp * 2) |>
  select(maxtemp, mintemp) |>
  head()
```

**summarize:**

We can use summarize(mean_mintemp = mean(mintemp)) to summarize the mintemp column as a mean
```{r}
weather |>
  summarize(mean_mintemp = mean(mintemp, na.rm = TRUE)) |>
  head()
```

**group_by:**

We can use group_by(location) to make the unit of observation location, and then use a function like summarize(avg_mintemp = mean(mintemp))
```{r}
weather |>
  group_by(location) |>
  summarize(avg_mintemp = mean(mintemp, na.rm = TRUE)) |>
  head()
```

**ungroup:**

You can reverse the unit of observation you select in group_by by using ungroup by, but only if you didn't use summarize. In this example, we use group_by(location), then mutate(avg_mintemp = mean(mintemp)), then ungroup()
```{r}
weather |>
  group_by(location) |>
  mutate(avg_mintemp = mean(mintemp, na.rm = TRUE)) |>
  ungroup() |>
  head()
```

**count:**

We have dataset hikes
```{r}
head(hikes)
```

We can use count(rating) to count the number of observations in the location column for each value
```{r}
hikes |>
  count(rating) |>
  head()
```
We can also use count() to count the total observations in the dataset
```{r}
hikes |>
  count() |>
  head()
```

**n:**

We can use n(rating) in a sumamrize function to count the number of observations in a group, similar to count
```{r}
hikes |>
  group_by(rating) |>
  summarize(total_trails = n(rating, na.rm = TRUE))
```
















