---
title: "Exam 2"
format: html
---

* this is a continuation of exam one

## Research question:

**What does the consumption of each food category in each country look like?**


## Import Necessary Data and Tools:

**Load in packages:**
```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
library(ggthemes)
```


**Load data:**
```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```


**Inspect Data:**
```{r}
str(fc)
```

```{r}
head(fc, 22)
```


**Inspect Country Variable:**
```{r}
fc |>
  group_by(country) |>
  summarize(sum(consumption)) |>
  str()
```

There are 130 observations, when you set the unit of observation to countries, so there are 130 countries in this dataset


**Inspect food_category variable:**
```{r}
fc |>
  distinct(food_category)
```

Here we have 11 distinct food categories, but it is better to rename them for the sake of making them easier to read in a viz

**Clean data:**
```{r}
fc_clean <- fc |>
  mutate(food_category = as.factor(food_category),
         food_category = fct_recode(food_category,
                                    "Lamb" = "Lamb & Goat",
                                    "Dairy" = "Milk - inc. cheese",
                                    "Wheat" = "Wheat and Wheat Products",
                                    "Nuts" = "Nuts inc. Peanut Butter"))
```

Double checking to make sure all variables are properly renamed
```{r}
fc_clean |>
  distinct(food_category)
```

## Sub-research Question 1: Which 5 countries consume the most food?

**Expected Dataframe:**
We want a dataframe in which countries are listed, followed by their total consumption


**Reshaping / Wrangling data:**
```{r}
country_consumption <- fc_clean |>
  group_by(country) |>
  summarize(total_consumption = sum(consumption))
  # pivot_wider(names_from = food_category, values_from = consumption) |>
  # mutate(total_consumption = Pork + Poultry + Beef + Lamb + Fish + Eggs + Dairy + Wheat + Rice + Soybeans + Nuts) |>
  # select(country, total_consumption)
```


**Visualizing Data:**
```{r}
country_consumption |>
  arrange(desc(total_consumption)) |>
  head(5) |>
  ggplot(aes(x = country, y = total_consumption)) +
    geom_col() +
    theme_solarized()
```


**Observations:**
Seemingly, the top consumers of food are exclusively European countries. Their geographic location varies from Eastern to Western, Baltics and Balkans and Scandinavian countries. They all consume roughly similar amounts of food, with Findland consuming the most.


## Sub-research Question 2: Which 5 countries consume the most of each food?

**Expected Dataframe:**
We want a dataframe where there is a variable for countries, a variable for the food type, and a variable for the consumption value. The food_category should correspond to the top 5 highest values in the consumption variable for its category. And the countries in the country variable will be the top 5 consumers of whatever food category is in the food variable.


**Reshaping / Wrangling Data:**
```{r}
top_food_consumers <- fc_clean |>
  select(-co2_emmission) |>
  group_by(food_category) |>
  slice_max(order_by = consumption, n = 5)
```


**Vizualizing Data:**
```{r}
ggplot(top_food_consumers, aes(x = food_category, y = consumption, fill = country)) +
  geom_col() +
  theme_solarized() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 5),
        axis.text.y = element_text(size = 5))
```


**Observations:**
Of all categories, dairy, wheat, and rice are the most consumed. The consumption of dairy, wheat, or rice is enough to outmatch the entire consumption of the top 5 countries combined in some categories like nuts or soybeans. What this means is that some of these categories are relatively easy to be a top 5 country in, because their consumption is very low anyway. 


## Sub-research Question 1: What does the consumption of each food look like?

**Expected Dataframe:**
We want a dataframe in which there is a country variable, a variable for food type, a corresponding variable for consumption, and a variable for the geometry of each country.


**Reshaping, Wrangling, and importing data:**

Find out which countries don't have a match in the geometry dataset
```{r}
fc_clean |>
  anti_join(ne_countries(), join_by(country == admin)) |>
  distinct(country)

st_set_geometry(ne_countries(), NULL) |>
  anti_join(fc_clean, join_by(admin == country)) |>
  distinct(admin)
```


Rename the countries so that they have a match and join the shapefile
```{r}
temp <- fc_clean |>
  select(-co2_emmission) |>
  group_by(food_category) |>
  mutate(consumption = (consumption - mean(consumption)) / sd(consumption)) |>
  mutate(country = fct_recode(country,
                              "United States of America" = "USA",
                              "Czechia" = "Czech Republic",
                              "Republic of Serbia" = "Serbia",
                              "eSwatini" = "Swaziland",
                              "North Macedonia" = "Macedonia",
                              "Mauritania" = "Mauritius",
                              "United Republic of Tanzania" = "Tanzania",
                              "Democratic Republic of the Congo" = "Congo",
                              "United Republic of Tanzania" = "Tanzania",
                              "The Bahamas" = "Bahamas",
                              "Taiwan" = "Taiwan. ROC"))

temp

global_consumption <- ne_countries() |>
  select(admin, geometry) |>
  left_join(temp, join_by(admin == country)) |>
  mutate(consumption = replace_na(consumption, 0)) |>
  pivot_wider(names_from = food_category, values_from = consumption) |>
  select(-`NA`) |>
  pivot_longer(cols = c(-geometry, -admin), names_to = "food_category", values_to = "consumption")

head(global_consumption)
```


**Vizualizing Data:**
```{r}
#| fig-height: 8
#| fig-width: 15
ggplot() +
  geom_sf(data = global_consumption,
          aes(fill = consumption)) +
  facet_wrap(~food_category) +
  coord_sf(crs = "+proj=robin") +
  theme_map() +
  theme(legend.position = c(0.8, 0.05))
```

**Observations:**
Around the world, dairy is heavily consumed asides from every part of the Asian continent (excluding Russia and Kazakhstan). There is missing data on much of Africa, but African countries with data also have low dairy consumption. All other food categories, in comparison to dairy, do not come close. The closest two categories are wheat (with high to moderate levels of consumption in Europe / Africa) and rice (with moderate levels of consumption in East, South, and Southeast Asia.















