---
title: "Solo Project"
format: html
---

## Research Question:

What is the distribution of Mormon Temples in the Contigious United States by State. (A Mormon Temple is not the same as a church. It holds particular religious importance)


## Step 1: Load Necessary Packages

```{r}
library(sf)
library(tidyverse)
library(dplyr)
```


## Step 2: Import / Clean / Wrangle the Data


**Import a shapefile of the United States and clean / wrangle it**

```{r}
messy_us_states_sf <- read_sf("../data/archive/cb_2018_us_state_500k.shp")

us_states_sf <- messy_us_states_sf |>
  filter(NAME != "Guam") |>
  filter(NAME != "Commonwealth of the Northern Mariana Islands") |>
  filter(NAME != "Hawaii") |>
  filter(NAME != "Alaska") |>
  filter(NAME != "Puerto Rico") |>
  filter(NAME != "United States Virgin Islands") |>
  filter(NAME != "American Samoa")
```


**Import data on mormon temples and clean / wrangle it**

```{r}
messy_mormon_temples <- read_csv("../data/ChurchofJesusChristTemples.csv")

mormon_temples_national <- messy_mormon_temples |>
  filter(Country == "United States") |>
  group_by(State) |>
  summarize(total_temples = n())
```


**Join the shapefile and the data on Mormon temples**

sidenote: make sure any NA values in the total_temples column become 0

```{r}
geospatial_mormon_temples <- us_states_sf |>
  left_join(mormon_temples_national, join_by(NAME == State)) |>
  mutate(total_temples = replace_na(total_temples, 0)) |>
  mutate(total_temples = cut(total_temples, 
                             breaks = c(0, 1, 6, 11, 16, 16-21, 30, Inf), 
                             labels = c("0", "1-5", "6-10", "11-15", "16-19", "20-29", "30+"), include.lowest = TRUE))

my_colors <- c(
  "0" = "#9E3D22",
  "1-5" = "#CA5621",
  "6-10" = "#E3732A",
  "11-15" = "#EFB27E",
  "16-20" = "#72A6CE",
  "20-29" = "#3A6B99",
  "30+" = "#2B5C8A")

```


## Step 3: Finally, create a choropleth map of the contiguous United States by total temples in each state

Note: this color palette uses a colorblind friendly palette

```{r}
#| fig-cap: "Choropleth map of the contigious U.S States colored by their total Mormon temples"
#| fig-alt: "A choropleth map of the contigious U.S states, colored by gradient from dark brown / orange to light blue and then to dark blue. The shows the number of Mormon temples each state has. Utah has the most temples, above 30, followed by California between 16 and 20. California is followed by Idaho, Texas, and Nevada at 11-15. Most other states have 6-10 or 1-5. A few, such as West Virginia, Alabamam, and many in New England have no temples. States on the West Coast or in the Rocky Mountains tend to have more Mormon temples than those in the South or Great Plains, but exceptions include Florida and Virginia in the South and Missouri in the Great Plains. There are around 1 temple per state in the Midwest, but Illinois, Michigan, and Ohio have more 6-10 temples Pennsylvania and New York have around 6-10 temples, but New Jersey and Maryland have 1-5 and Delaware has 0."

ggplot() +
  geom_sf(data = geospatial_mormon_temples,
          aes(fill = total_temples)) +
  scale_fill_manual(name = "Total Mormon Temples", values = my_colors) +
  theme_void() +
  coord_sf(crs = 5070)
```

Interesting. Utah has an incredibly high number of temples. I wonder how temples are distributed spatially in Utah.

## Research Question 2: 

What is the distribution of Mormon temples in Utah by County


## Step 1: Load Necessary Packages

Packages were already loaded earlier


## Step 2: Import / Clean / Wrangle the Data


**Use the already imported data on Mormon temples, and clean / wrangle it**

```{r}
mormon_temples_utah <- messy_mormon_temples |>
  filter(Country == "United States") |>
  filter(State == "Utah") |>
  group_by(County) |>
  summarize(total_temples = n())
```


**Import data of counties' fips codesm and their names, across the United States and clean / wrangle it**

```{r}
utah_county_fips <- read_csv("../data/state_and_county_fips_master.csv")
```


**Import a shapefile of counties in Utah, and clean / wrangle it while also joining the dataset on county fips codes**

```{r}
messy_utah_counties_sf <- read_sf("../data/UtahCountyBoundaries_6245553605277884471/Counties.shp")

utah_counties_sf <- messy_utah_counties_sf |>
  mutate(FIPS_STR = as.numeric(FIPS_STR)) |>
  left_join(utah_county_fips, join_by(FIPS_STR == fips)) |>
  mutate(name = str_replace(name, " County", ""))
```


**Join the data on Mormon temples to the shapefile**

sidenote: make sure any NA values in the total_temples column become 0. Also, create classes and a color palette to use for them in the choropleth map.

```{r}
mormon_temples_utah_sf <- utah_counties_sf |>
  left_join(mormon_temples_utah, join_by(name == County)) |>
  mutate(total_temples = replace_na(total_temples, 0)) |>
  mutate(total_temples = cut(total_temples, 
                             breaks = c(0, 1, 2, 4, 6, 8, Inf), 
                             labels = c("0", "1", "2-3", "4-5", "6-7", "8+"), include.lowest = TRUE))

my_colors <- c(
  "0" = "#F4D166",
  "1" = "#BEC45C",
  "2-3" = "#6EAC56",
  "4-5" = "#38884C",
  "8+" = "#146C36")
```


## Step 3: Create a choropleth map of Utah's counties by total temples in each state


```{r}
#| fig-cap: "Choropleth map of Utah's counties and how many Mormon temples they have"
#| fig-alt: "A choropleth map of Utah's counties, using a gradient from yellow to dark green, shows the number of Mormon temples each county has. Most of the counties in the state have 0 temples. There is a cluster of counties near the center of the state that have temples. Disconnected from the central cluster, there are 2 other counties with one temple in them, one in the north and one in the southwest. The central cluster has dark green counties, most of them having more than one temple and one county with 8 or more temples. In the cluster, a pattern emerges where southern counties have more temples than those in the north, with the southernmost county that has 1 temple being the exception as it borders the most temple rich county to its north."

ggplot() +
  geom_sf(data = mormon_temples_utah_sf,
          aes(fill = total_temples)) +
  scale_fill_manual(name = "Total Mormon Temples", values = my_colors) +
  theme_void()
```

Interestingly, it seems as though nearly all of the Mormon temples focused themselves into one cluser of counties near the center of the state. There are only 2 other counties that are disconnected from the cluster of counties, one in the north and one in the southwest. Most counties, across the state, have no temples. In the central counties, there is a gradient from the south to north, with a decreasing number of temples in each county, with the exception of the southernmost county which has only 1 temple.


## Conclusion:

Throughout the contigious United States, the states with the most Mormon temples tend to be in the Rocky Mountains or on the West Coast. The region of states with the fewest temples is the Great Plains. But Utah stands out as the state with the highest concentration of temples, far surpassing all other states. When you investigate Utah's spatial distribution of temples by county, you discover that they are focused in a cluster of counties in the center of the state, with 2 other outlier counties (that have only one temple) in the state. The counties clustered in the center have an incredibly dense number of temples, in comparison to the rest of the united cities. One singular county has 8+ temples, more temples than most states have. This leads to the second conclusion about the distribution of Mormon temples in the contigious United States: Utah has an incredibly spatially concentrated distribution of temples near the center of its state, making the state surpass the number of temples per state across the contigious United States






